---
title: "MCreproduce"
output: pdf_document
date: "2026-03-09"
---

1. First, learn how to do a Monte-Carlo simulation.
2. Then figure out how to implement the multi-outcomes estimator.
3. Then generate the data.
4. Then run the simulation.


# Monte-Carlo Simluation Study in R

- True value of target parameter is known and is used to generate the data.

What is the target parameter?
- The target parameter is the counterfactual estimate under control.

MC simulation framework:
1. Generate $S$ independent data sets under the conditions of interest
2. Compute the statistic of interest (bias) for each one.
3. Mean of the biases should be a good estimator for true bias.

Before writing code:
- Bias is the estimator of interest.
- Fully specify data generating model
- Random number generator

# Toy Example of Monte-Carlo Simulation Study

\textcolor{red}{Have I properly set the seed for reproducibility?}

We propose the sample mean as an estimator of the population mean. Assume data are generated from $N(0,1)$. We find the bias of our estimator.

```{r}
# for reproducibility
set.seed(215)

# a function to generate S independent data sets, reproducibly
mc_experiment <- function(S, n) {
  # list of what we will return
  return_vals <- list()
  # will be populated with multiple independent data sets
  data_list <- list()
  # number of data sets is S
  for (s in 1:S) {
    # number of normal (0,1) samples is n
    data <- rnorm(n)
    data_list[[s]] <- data
  }
  # now we calculate the sample means
  sample_means <- list()
  for (s in 1:S) {
    X_bar <- mean(data_list[[s]])
    sample_means[[s]] <- X_bar
  }
  # calculate the MC estimator of the bias (S replicates)
  sample_means <- unlist(sample_means)
  mc_bias <- mean(sample_means)
  # put the values to return in a list and return them
  return_vals[["data_list"]] <- data_list
  return_vals[["mc_bias"]] <- mc_bias
  return (return_vals)
}
```

```{r}
# to have reproducible results, these have to be the same
# changing these experiment hyperparameters changes the results.
num_experiments = 3
S = 2
n_obs = 10

# run the experiment - we get the same thing every time.
for (e in 1:num_experiments) {
  # get a list of data sets
  return_vals <- mc_experiment(S, n_obs)
  data_list <- return_vals[["data_list"]]
  mc_bias <- return_vals[["mc_bias"]]
  
  # print the generated data sets
  for (s in 1:S) {
    cat("data_list", s, ": ", data_list[[s]],"\n")
  }
  cat("mc_bias: ",mc_bias,"\n")
}
```

```{r}
# using micro benchmark to time mc_experiment
iterate <- function(S, n_obs) {
  mc_experiment(S,n_obs)
}

library(microbenchmark)
microbenchmark(
  iterate(S,n_obs)
)
```

# Improved MC Simulation Study Using Mclapply

```{r}



# a function to generate S independent data sets, reproducibly
mc_experiment <- function(S) {
  # hard code the size of the observation vector
  n <- 10
  # list of what we will return
  return_vals <- list()
  # will be populated with multiple independent data sets
  data_list <- list()
  # number of data sets is S
  for (s in 1:S) {
    # number of normal (0,1) samples is n
    data <- rnorm(n)
    data_list[[s]] <- data
  }
  # now we calculate the sample means
  sample_means <- list()
  for (s in 1:S) {
    X_bar <- mean(data_list[[s]])
    sample_means[[s]] <- X_bar
  }
  # calculate the MC estimator of the bias (S replicates)
  sample_means <- unlist(sample_means)
  mc_bias <- mean(sample_means)
  # put the values to return in a list and return them
  return_vals[["data_list"]] <- data_list
  return_vals[["mc_bias"]] <- mc_bias
  return (return_vals)
}
```

```{r}
num_cores = 14

iterate <- function(K,num_cores) {
  output <- mclapply(1:K, function(i) mc_experiment(K), mc.cores=num_cores)
  return (output)
}
```

```{r}
microbenchmark(
  for (i in 1:5){
    iterate(1000,14)
  }
)
```