---
title: "McSimDatGen"
output:
  pdf_document: default
  html_document: default
---
\section{Generating Data}

Goal is to generate outcomes using the model
$$
Y_{it,k}^0 = \delta_{t,k} + \mathbf{Z}_{i,t}' \boldsymbol{\theta}_{k} + \boldsymbol{\mu}_{i,t}\boldsymbol{\lambda}_{k} ,\quad\text{leaving out contribution from }\varepsilon_{itk}.
$$ which has time invariant factor loadings.
In particular:
\begin{itemize}
\item For each pair $(t,k)$, $\delta_{t,k}$ is a scalar. 
\item For each pair $(i,t)$, the vector $\mathbf{Z}_{i,t}$ will have dimension $(\texttt{n\_covariates},1)$. (Observed covariate measurements)
\item For each $k$, the vector $\boldsymbol{\theta}_{k}$ will have dimension $(\texttt{n\_covariates},1)$. 
\item For each pair $(i,t)$, the vector $\boldsymbol{\mu}_{i,t}$ will have dimension $(\texttt{latent\_dim},1)$. (Latent factors)
\item For each $k$, the vector $\boldsymbol{\lambda}_{k}$ will have dimension $(\texttt{latent\_dim},1)$
\end{itemize}


\subsection{Generate $\delta_{t,k}$}

For $k \in \{1, \ldots, \texttt{n\_outcomes}\}$, we generate
$$
\delta_{t,k} = \delta_{t-1,k} + \xi, \quad \xi\sim N(0,0.1), \delta_{1,k} \sim N(1,1).
$$
The first row is outcome $k=1$, second is $k=2$, etc. After some treatment time, we change $\xi$ to be drawn from $N(1,0.1)$, representing a treatment effect.

\subsection{Generating the $\mathbf{Z}_{i,t}$}

Let $\mathbf{Z}_{1,1} = (\varphi_1, \varphi_2,\varphi_3)$, and $\zeta_1 = (\zeta_{11}, \zeta_{12},\zeta_{13})$. We set:
$$
\mathbf{Z}_{1,2} = \mathbf{Z}_{1,1} + \zeta_1,
$$ and 
$$
\mathbf{Z}_{1,3} = \mathbf{Z}_{1,1} + 2 \zeta_1,
$$
to give time trends to the covariates.

\subsection{Generate the $\boldsymbol{\mu}_{i,t}$}

We use a similar procedure to the generation of the observed covariates. For instance, the vector $\boldsymbol{\mu}_{1,1}$ has dimension 2, and we add a latent time trend in each latent factor through the repeated addition of $\zeta_{1}*$ (star to differentiate from $\zeta_1$). The three matrices we end up with correspond to the three units in the simulation.

\subsection{Generate $\boldsymbol{\theta}_{k}$ and $\boldsymbol{\lambda}_{k}$}
Create the $k$ loading vectors $\boldsymbol{\theta}_{k}$ each with dimension $(\texttt{n\_covariates},1)$.

Create the $k$ loading vectors for $\boldsymbol{\lambda}_{k}$ each with dimension $(\texttt{latent\_dim},1)$.

```{r, echo=FALSE}
# Set up
library(MASS)
set.seed(215)

### ------------------------------------------------------------ ###

# t = n_timepoints
# k = n_outcomes

# a function that creates the delta tk entries.
make_dtk <- function(n_timepoints, n_outcomes, treatment_effect, treatment_time) {
  # initialize an empty matrix of dim t,k
  dtk_mat <- matrix(NA, n_timepoints, n_outcomes)
  # first row is random noise of length n_outcomes
  dtk_mat[1,] <- rnorm(n_outcomes, 1, 1)
  # fill in the remaining rows:
  for (i in 2:n_timepoints) {
    if (treatment_effect==TRUE & i > treatment_time) {
      effect <- (i-treatment_time) * rnorm(n_outcomes,1,0.1)
    } else {
      effect <- rnorm(n_outcomes,0,0.1)
    }
    prev <- if (i==1) dtk_mat[1,] else dtk_mat[i-1,]
    dtk_mat[i,] <- prev + effect
  }
  return(dtk_mat)
}

### ------------------------------------------------------------ ###

make_Z_matrix <- function(z, zeta, n_timepoints) {
  z = z
  zeta = zeta
  n_timepoints = n_timepoints
  
  mat <- matrix(NA, nrow=n_timepoints, ncol=length(z))
  for (i in 1:n_timepoints) {
    mat[i,] <- z + ((i-1) * zeta)
  }
  return(mat)
}

### ------------------------------------------------------------ ###

# a function to make the time varying covariates
# NB: number of predictors is dimension of covariate vectors. 
# Not necessarily the same as the length of the outcome vectors
make_Z_matrices <- function(n_predictors, n_timepoints) {
  n_predictors = n_predictors
  n_timepoints = n_timepoints
  
  mean = rep(0, n_predictors)
  cov = diag(n_predictors) # identity matrix
  
  # generating initial covariate measurements
  # each row represents initial covariate measurements at t = 1, for each i in 1,...,n_units
  Z <- mvrnorm(n = n_units, mu = mean, Sigma = cov) 
  # Zeta determines how covariates change over time.
  Zeta <- mvrnorm(n = n_units, mu = mean, Sigma = cov) #* 0.1 # to smooth time trends
  
  # a function to apply Zeta to each row of the initial Z matrix
  
  # columns of Z_matrices[[1]], eg, represent a covariate changing over time.
  Z_matrices <- lapply(1:n_units, function(i) {
    make_Z_matrix(Z[i,], Zeta[i,], n_timepoints = n_timepoints)
  })
  return(Z_matrices)
}

### ------------------------------------------------------------ ###

# makes the time varying latent factors.
make_M_matrices <- function(latent_dim, n_timepoints) {
  latent_dim = latent_dim
  n_timepoints = n_timepoints
  
  mu_mean <- rep(0, latent_dim)
  mu_cov  <- diag(latent_dim)
  
  # Initial factor values for each unit
  M <- mvrnorm(n = n_units, mu = mu_mean, Sigma = mu_cov)
  
  # Factor drift over time (analogous to Zeta)
  Xi <- mvrnorm(n = n_units, mu = mu_mean, Sigma = mu_cov) #*0.1 # to smooth time trends
  
  M_matrices <- lapply(1:n_units, function(i) {
    make_Z_matrix(M[i,], Xi[i,], n_timepoints = n_timepoints)
  })
  return(M_matrices)
}

### ------------------------------------------------------------ ###

# make the time invariant covariate loadings
make_theta <- function(n_outcomes, n_predictors) {
  n_outcomes = n_outcomes
  n_predictors = n_predictors
  
  theta <- matrix(NA, nrow = n_outcomes, ncol = n_predictors)
  for (i in 1:n_outcomes) {
    theta[i,] <- rnorm(n=n_predictors, mean=0, sd=0.1/sqrt(n_predictors))
  }
  if (verbose==TRUE) {
    cat("Loadings for observed covariates: \n")
    cat("(One row is one theta_k vector) \n")
    print(theta)
    cat("\n")
  }
  return (theta)
}

### ------------------------------------------------------------ ###

# generate time invariant latent factor loadings
make_lambda <- function(n_outcomes, latent_dim) {
  n_outcomes = n_outcomes
  latent_dim = latent_dim
  
  lambda <- matrix(NA, nrow = n_outcomes, ncol = latent_dim)
  for (i in 1:n_outcomes) {
    lambda[i,] <- rnorm(n=latent_dim, mean=0, sd=0.1/sqrt(latent_dim))
  }
  
  if (verbose==TRUE) {
    cat("Loadings for unobserved latent factors: \n")
    cat("(One row is one lambda_k vector) \n")
    print(lambda)
    cat("\n")
  }
  return(lambda)
}

### ------------------------------------------------------------ ###

# standardize per outcome using only the pretreatment info.
scale_function <- function(Y,n_units, n_timepoints,n_outcomes,treatment_time) {
  for (i in 1:n_outcomes) {
    mat <- matrix(Y[,,i], nrow=n_units, ncol = n_timepoints)
    observable_mat <- mat[,1:treatment_time]
  
    # calculate the mean and sd of the observable data matrix
    tot_mean <- mean(observable_mat)
    vect <- as.vector(observable_mat)
    tot_sd <- sd(vect)
    
    mat <- (mat - tot_mean) / tot_sd
    Y[,,i] <- mat
  }
  Y_scaled <- Y
  return (Y_scaled)
}

### ------------------------------------------------------------ ###

# notes for (n_units, n_timepoints)
# (4, 10) gives good similarity (4,20) is even better
# (10, 10) gives great similarity (10, 20) is also extremely good
# turning up n_outcomes to 3 it still works great with (10,20) and (10,10)
# (8,8) embedding into 9 is a good example to show jing

# inputs to the generate data function
n_units = 3 # number of units (i)
n_timepoints = 10 # number of time points (t)
n_outcomes = 2 # number of outcomes (k)

n_predictors = 10 # number of predictors (length of each vector z_{i,t}) (p)
latent_dim = 4 # dimension of latent space. (l)

treatment_effect = FALSE # should we generate the data with a treatment effect
treatment_time = 5 # must be less than n_timepoints

verbose = FALSE 

### ------------------------------------------------------------ ###

# generates the synthetic data. Calls several other functions, above
generate_data <- function(n_units, n_timepoints, n_outcomes, n_predictors, latent_dim, treatment_effect, treatment_time, verbose) {
  dtk_matrix <- make_dtk(n_timepoints,n_outcomes, treatment_effect, treatment_time) * 0.1
  if (verbose == TRUE) {
    cat("dtk_matrix: \n")
    print(dtk_matrix)
    cat("\n")
  }
  
  Z_matrices <- make_Z_matrices(n_predictors, n_timepoints)
  M_matrices <- make_M_matrices(latent_dim, n_timepoints)
  theta <- make_theta(n_outcomes, n_predictors)
  lambda <- make_lambda(n_outcomes, latent_dim)
  
  # an array to store the synthetic outcomes
  Y <- array(NA, dim = c(n_units, n_timepoints, n_outcomes))
  
  # add everything together
  for (i in 1:n_units) {
    for (k in 1:n_outcomes) {
      for (t in 1:n_timepoints) {
        delta_tk <- dtk_matrix[t,k]
        # take the product of n(0,1) * n(0,1) has variance 1. We add together n_predictors. 
        # so the variance of the product is n_predictors. We standardize when we are generating
        # the loadings /sqrt(n_predictors)
        z_theta_itk <- (Z_matrices[[i]][t,] %*% theta[k,]) 
        # also standardized the same way here
        mu_lambda_itk <- (M_matrices[[i]][t,] %*% lambda[k,]) 
        
        Y[i,t,k] <- delta_tk + z_theta_itk + mu_lambda_itk
      }
    }
  }
  if (verbose==TRUE) {
    cat("The final data Y, by unit: \n")
    for (i in 1:n_units) {
      cat("Unit ",i," timepoints x outcomes\n")
      print(Y[i,,])
      cat("\n")
    }
  }
  
  
  
  Y <- scale_function(Y,n_units, n_timepoints,n_outcomes,treatment_time)
  
  # stack the matrices                                    
  Y_final <- do.call(rbind, lapply(1:n_timepoints, function(i) t(Y[,i,])))
  cat("All unit outcomes stacked: \n")
  print(Y_final)
  cat("\n")
  
  # output information
  write.csv(Y_final, "outcomes.csv", row.names = FALSE)
  metadata <- c(n_units, n_timepoints, n_outcomes)
  write.csv(metadata, "metadata.csv", row.names = c("n_units","n_timepoints","n_outcomes"))
}

### ------------------------------------------------------------ ###

# call the function
generate_data(n_units, n_timepoints, n_outcomes, n_predictors, latent_dim, treatment_effect, treatment_time, verbose)


```


```{r,echo=FALSE}

# # Staging ground for standardizing.
# 
# # need to normalize using only pre-treatment - first t_0 rows
# set.seed(215)
# 
# n_units = 3
# n_timepoints = 3
# n_outcomes = 2
# treatment_time = 2
# 
# # Creates a 3x4x2 array of random integers between 1 and 100
# X <- array(sample(1:100, 24, replace = TRUE), dim = c(n_units, n_timepoints, n_outcomes))
# print(X)
# 
# for (i in 1:n_outcomes) {
#   mat <- matrix(X[,,i], nrow=n_units, ncol = n_timepoints)
#   cat("pre-treatment sub-matrix: \n")
#   observable_mat <- mat[,1:treatment_time]
#   print(observable_mat)
# 
#   # calculate the mean and sd of the observable data matrix
#   tot_mean <- mean(observable_mat)
#   vect <- as.vector(observable_mat)
#   tot_sd <- sd(vect)
#   
#   mat <- (mat - tot_mean) / tot_sd
#   X[,,i] <- mat
# }
```

# Eli Ben Michael's code
```{r}

######-------------------------------------------------------------- ELI BEN-MICHAEL'S CODE ---------------------------------------------------#######

# ## Install devtools if not already installed
# install.packages("devtools", repos='http://cran.us.r-project.org')
# library(devtools)
# ## Install augsynth from github
# devtools::install_github("ebenmichael/augsynth", force = T)
# library(augsynth)
## general libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(MASS)
rm(list = ls())
# The palette with black:
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# Change the weights in model to reflect importance of common factors
rho <- 1
bias_fig_name <- 'figure/all_bias_sims_n50_rho1.pdf'
imbalance_fig_name <- 'figure/all_imbalance_sims_n50_rho1.pdf'

# rho <- 0.5
# bias_fig_name <- 'figure/all_bias_sims_n50_rho05.pdf'
# imbalance_fig_name <- 'figure/all_imbalance_sims_n50_rho05.pdf'

# rho <- 0
# bias_fig_name <- 'figure/all_bias_sims_n50_rho0.pdf'
# imbalance_fig_name <- 'figure/all_imbalance_sims_n50_rho0.pdf'

dir.create("figure", showWarnings = FALSE)

##### SCM function -------------------------------------

#' Solve the synth QP directly
#' @param X1 Target vector
#' @param X0 Matrix of control outcomes
#' @noRd
synth_qp <- function(X1, X0) {
  
  Pmat <- X0 %*% t(X0)
  qvec <- - t(X1) %*% t(X0)
  
  n0 <- nrow(X0)
  A <- rbind(rep(1, n0), diag(n0))
  l <- c(1, numeric(n0))
  u <- c(1, rep(1, n0))
  
  settings = osqp::osqpSettings(verbose = FALSE,
                                eps_rel = 1e-8,
                                eps_abs = 1e-8)
  sol <- osqp::solve_osqp(P = Pmat, q = qvec,
                          A = A, l = l, u = u, 
                          pars = settings)
  
  return(sol$x)
}

##### Bias 
SCMbias <- function(Y0,Y1,weights) {
  synthY0 <- (Y0%*%weights)
  gap <- Y1-synthY0 
  return(gap)
}

#### RMSE
rmse <- function(Y0,Y1,weights) {
  synthY0 <- (Y0%*%weights)
  gap <- Y1-synthY0 
  return(sqrt(mean(gap^2)))
}

#' Given the factors and factor loadings, simulate the data
#' @param factors Factors
#' @param mu Factor loadings
#' @param n number of units
#' @param trt treatment indicator vector
#' @param k_total number of outcomes
#' @param t_total number of pre-treatment periods
#' @param variance variance of the noise term
factor_sim <- function(model,n,trt,k_total,t_total,variance) {
  
  ## generate factor model data
  
  epsilon <- matrix(rnorm((n * t_total) * k_total, sd = sqrt(variance)), ncol = k_total)

  out <- model[1:(n * t_total),] + epsilon # (NXT) x (K) matrix of outcomes
  t_trt <- rep(trt,t_total) # select the periods for the treated unit
  out_trt <- matrix(out[which(t_trt==1),],nrow = t_total,ncol =k_total)  # T x K matrix of outcomes for the treated unit
  out_control <- out[which(t_trt==0),] # ((N-1)xT) x (K) matrix of outcomes for the control units
  
  
  ## calculate SCM weights
  out_trt_sep <- out_trt[,1]
  out_control_sep <- matrix(out_control[,1], nrow = n-1,ncol=t_total)
  w_sep <- synth_qp(out_trt_sep, out_control_sep);
  r.svd <- svd(rbind(out_control_sep,out_trt_sep))
  largest_svd <- r.svd$d[1]^2/sum(r.svd$d^2)
  cond <-  r.svd$d[1]/r.svd$d[t_total]
 
  
  out_trt_cat <- matrix(out_trt, nrow = t_total*k_total, ncol = 1)
  out_control_cat <- matrix(out_control, nrow = n-1,ncol=t_total*k_total)
  w_cat <- synth_qp(out_trt_cat, out_control_cat);
  r.svd <- svd(rbind(out_control_cat,t(out_trt_cat)))
  largest_svd <- cbind(largest_svd, r.svd$d[1]^2/sum(r.svd$d^2))
  cond <- cbind(cond,  r.svd$d[1]/r.svd$d[length(r.svd$d)])
  
  out_trt_avg <- rowMeans(out_trt)
  out_control_avg <- matrix(rowMeans(out_control), nrow = n-1,ncol=t_total)
  w_avg <- synth_qp(out_trt_avg, out_control_avg);
  r.svd <- svd(rbind(out_control_avg,t(out_trt_avg)))
  largest_svd <- cbind(largest_svd, r.svd$d[1]^2/sum(r.svd$d^2))
  cond <- cbind(cond,  r.svd$d[1]/r.svd$d[t_total])
 
  
  ## calculate bias
  model_t1 <- model[((n * t_total)+1):(n * (t_total+1)),]
  oracle_bias_sep <- SCMbias(model_t1[-trt,1],model_t1[trt,1],w_sep)
  oracle_bias_cat <- SCMbias(model_t1[-trt,1],model_t1[trt,1],w_cat)
  oracle_bias_avg <- SCMbias(model_t1[-trt,1],model_t1[trt,1],w_avg)
  
  ## calculate pre-treatment imbalance
  imbalance_sep <- rmse(t(out_control_sep),out_trt_sep,w_sep)
  imbalance_cat <- rmse(t(out_control_cat),out_trt_cat,w_cat)
  imbalance_avg <- rmse(t(out_control_avg),out_trt_avg,w_avg)
  
  # return(list("oracle_bias" = c(oracle_bias_sep,oracle_bias_cat,oracle_bias_avg),
  #             "imbalance" = c(imbalance_sep,imbalance_cat,imbalance_avg),
  #             "largest_svd"=largest_svd,"cond"=cond ) )
  
  ##### Switching this for the above (JS) #######
  return(list("oracle_bias" = c(oracle_bias_sep, oracle_bias_cat, oracle_bias_avg),
              "imbalance" = c(imbalance_sep, imbalance_cat, imbalance_avg),
              "largest_svd" = largest_svd,
              "cond" = cond,
              "out" = out,           # (NxT) x K matrix, noisy observed outcomes
              "out_trt" = out_trt,   # T x K matrix for treated unit
              "out_control" = out_control  # ((N-1)xT) x K matrix for controls
  ))
}


# factor loadings
n <- 50 # number of units, the first unit is treated
# mu <- c(2,1,3:n); # hand-code factor loadings for n = 5
mu <- list(
  c = seq(5, 1, length.out = n) # N x 1 vector
)
mu$c[c(1, 2)] <- mu$c[c(2, 1)]; # hand-code factor loadings for n = 50 (common part)

trt <- numeric(n); trt[1] <- 1; # select the unit with the second largest loadings to be the treated unit

settings_list <- list(
  list(t_total=10, k_total=4, file_suffix="T_0=10,K=4"),
  list(t_total=10, k_total=10, file_suffix="T_0=10,K=10"),
  list(t_total=40, k_total=4, file_suffix="T_0=40,K=4"),
  list(t_total=40, k_total=10, file_suffix="T_0=40,K=10")
)     # incorporate all settings into a loop

all_bias_sim_long <- data.frame()
all_imbalance_sim_long <- data.frame()
all_largest_svd_sim_long <- data.frame()
all_cond_sim_long <- data.frame()

# ADDING THIS TO EXTRACT FACTOR MODEL DATA (JS)
example_data <- list()

for (setting in settings_list) {
  t_total <- setting$t_total
  k_total <- setting$k_total   
  
  factors <- list(
    c = seq(0.5, 1, length.out = t_total+1) # hand-code factor (common part)
  )
  factors$i <- matrix(0,nrow=t_total+1, ncol = k_total) # outcome idiosyncratic part of factor
  set.seed(1)
  for ( k in 1:k_total ) {
    series <- arima.sim(model = list(ar = 0.5), n = t_total+1) # independent AR series across outcomes
    factors$i[,k] <- 0.5 + (series - min(series)) * (1 - 0.5) / (max(series) - min(series))  # rescale so factors are in same range
    factors$i[,k] <- sort(factors$i[,k])
  }
  
  set.seed(1)
  mu$i <- matrix(0,nrow=n, ncol = k_total) # outcome idiosyncratic part of factor
  for (k in (1:k_total)) {
    mu$i[,k] <- rnorm(n)
    mu$i[,k] <- 1 + mu$i[,k] * (5 - 1) / (max(mu$i[,k] ) - min(mu$i[,k] )) # rescale so bias is in same range
  }
  cor(mu$i[,1],mu$c)
  cor(mu$i[,1],mu$i[,2])
  
  theo_w <- synth_qp(mu$c[trt], as.matrix(mu$c[-trt])); # confirm we can find the oracle weights
  SCMbias(mu$c[-trt]*factors$c[t_total+1], mu$c[trt]*factors$c[t_total+1],theo_w)
  
   
  mu$i[trt,] <- t(theo_w)%*%mu$i[-trt,] # overwrite the idiosyncratic loadings to ensure oracle weights exist
  
  theo_w <- synth_qp(c(mu$c[trt],mu$i[trt,]), as.matrix(cbind(mu$c[-trt],mu$i[-trt,]))); # confirm we can find the oracle weights
  SCMbias(mu$c[-trt]*factors$c[t_total+1] + mu$i[-trt,1]*factors$i[t_total+1,1],
          mu$c[trt]*factors$c[t_total+1] + mu$i[trt,1]*factors$i[t_total+1,1], 
          theo_w)
  
  # create the model components
  factors$i[,1] <- factors$c
  # Repeat each row of factors N times (expand to NT x K)
  factors_expanded <- factors$i[rep(1:(t_total+1), each = n), ]
  mu$i[,1] <- mu$c
  # Repeat mu T times (expand to NT x K)
  mu_expanded <- mu$i[rep(1:n, (t_total+1)), ]
  
  # Element-wise multiplication
  model <- rho * as.vector(factors$c %x% mu$c) + (1-rho) * factors_expanded * mu_expanded
  # simulation
  
  
  n_sim <- 1000
  columns <- c("sep","cat","avg") 
  bias_sim <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
  colnames(bias_sim) <- columns
  
  imbalance_sim <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
  colnames(imbalance_sim) <- columns
  
  largest_svd_sim <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
  colnames(largest_svd_sim) <- columns 
  cond_sim <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
  colnames(cond_sim) <- columns 
  
  for (s in 1:n_sim) {
    set.seed(s)
    result <- factor_sim(model,n,trt,k_total,t_total,variance=1)
    # result <- factor_sim(mu,factors_tk,n,trt,k_total,t_total,variance=1)
    bias_sim[s,] <- result$oracle_bias
    imbalance_sim[s,] <- result$imbalance
    largest_svd_sim[s,] <- result$largest_svd
    cond_sim[s,] <- result$cond
    
    ###### Adding this (JS) #####
    if (s == 1) {
      example_data[[setting$file_suffix]] <- list(
        out = result$out,
        out_trt = result$out_trt,
        out_control = result$out_control
      )
    }
    #############################
  }
  
  # check the results
  summary(bias_sim)
  summary(imbalance_sim)
  colMeans(largest_svd_sim)
  colMeans(cond_sim)
  
  
  bias_sim_long <- bias_sim %>% 
    pivot_longer(columns, names_to = "method", values_to = "bias") %>% mutate(setting = setting$file_suffix)
  imbalance_sim_long <- imbalance_sim %>% 
    pivot_longer(columns, names_to = "method", values_to = "imbalance") %>% mutate(setting = setting$file_suffix)
  
  all_bias_sim_long <- rbind(all_bias_sim_long, bias_sim_long)
  all_imbalance_sim_long <- rbind(all_imbalance_sim_long, imbalance_sim_long)
  
  largest_svd_sim_long <- largest_svd_sim %>% 
    pivot_longer(columns, names_to = "method", values_to = "svd") %>% mutate(setting = setting$file_suffix)
  cond_sim_long <- cond_sim %>% 
    pivot_longer(columns, names_to = "method", values_to = "cond") %>% mutate(setting = setting$file_suffix)
  
  all_largest_svd_sim_long <- rbind(all_largest_svd_sim_long, largest_svd_sim_long)
  all_cond_sim_long <- rbind(all_cond_sim_long, cond_sim_long)
}

# Check the factor structure based on the share of variation explained and the conditional numbers
all_largest_svd_sim_long %>% filter(setting == "T_0=40,K=10") %>% group_by(method) %>%
  summarize(mean_x = mean(svd))

ratio <- all_cond_sim_long %>% filter(setting == "T_0=40,K=10" & method == "avg") %>% dplyr::select(cond) 
ratio <- ratio / all_cond_sim_long %>% filter( setting == "T_0=40,K=10" & method == "sep") %>% dplyr::select(cond) 
avg_cond_avg_v_sep <- mean(t(ratio)-1)
cat("Average ratio of conditional number for average over separate SCM-1",avg_cond_avg_v_sep,"\n" )
# Adjust the order of the setting
all_bias_sim_long$setting <- factor(all_bias_sim_long$setting, levels = c("T_0=10,K=4", "T_0=10,K=10", "T_0=40,K=4", "T_0=40,K=10"))
all_imbalance_sim_long$setting <- factor(all_imbalance_sim_long$setting, levels = c("T_0=10,K=4", "T_0=10,K=10", "T_0=40,K=4" , "T_0=40,K=10"))


# pdf('figure/all_bias_sims_n50_p05.pdf') #n09 = -0.9 AR; p09 = +0.9
pdf(bias_fig_name)
all_bias_sim_long %>%
  ggplot(aes(x=method, y=bias, fill=method)) +
  geom_boxplot(notch=FALSE,outlier.shape=NA) +
  stat_summary(fun=mean, geom="point", shape=20, size=2, color="black", fill="black") +
  facet_wrap(~setting, scales = "fixed", nrow = 2) + # Set scales as fixed
  geom_hline(yintercept = 0, color = "black") + # Add a horizontal line at 0 to all facets
  ylim(-0.25,2) + 
  labs(y = "Bias", x = "SC Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),
        axis.text.y = element_text(size=12),
        axis.title.x = element_text(size=14, face="bold"),
        axis.title.y = element_text(size=14, face="bold"),
        plot.title = element_text(size=16, face="bold", hjust=0.5),
        plot.margin = unit(c(0,0,0,0), "lines"),
        aspect.ratio = 3/4) +
  scale_fill_manual(values=cbp1)
dev.off()

pdf(imbalance_fig_name)
all_imbalance_sim_long %>%
  ggplot(aes(x=method, y=imbalance, fill=method)) +
  geom_boxplot(notch=FALSE,outlier.shape=NA) +
  stat_summary(fun=mean, geom="point", shape=20, size=2, color="black", fill="black") +
  facet_wrap(~setting, scales = "fixed", nrow = 2) + 
  geom_hline(yintercept = 0, color = "black") + # Add a horizontal line at 0 to all facets
  ylim(0,1.75) + 
  labs(y = "Imbalance", x = "SC Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size =12),
        axis.text.y = element_text(size=12),
        axis.title.x = element_text(size=14, face="bold"),
        axis.title.y = element_text(size=14, face="bold"),
        plot.title = element_text(size=16, face="bold", hjust=0.5),
        plot.margin = unit(c(0,0,0,0), "lines"),
        aspect.ratio = 3/4)+
  scale_fill_manual(values=cbp1)
dev.off()

```

```{r}
# Extracting the data from the factor model.
example_data[["T_0=10,K=4"]]$out
example_data[["T_0=40,K=10"]]$out_trt

treated_data <- example_data[["T_0=40,K=10"]]$out_trt
control_data <- example_data[["T_0=40,K=10"]]$out_control
```

