---
title: "McSimDatGen"
output:
  pdf_document: default
  html_document: default
---
\section{Generating Data}

Goal is to generate outcomes using the model
$$
Y_{it,k}^0 = \delta_{t,k} + \mathbf{Z}_{i,t}' \boldsymbol{\theta}_{t,k} + \boldsymbol{\mu}_{i,t}\boldsymbol{\lambda}_{t,k} ,\quad\text{leaving out contribution from }\varepsilon_{itk}.
$$
In particular:
\begin{itemize}
\item For each pair $(t,k)$, $\delta_{t,k}$ is a scalar. 
\item For each pair $(i,t)$, the vector $\mathbf{Z}_{i,t}$ will have dimension $(3,1)$. (Observed covariate measurements)
\item For each pair $(t,k)$, the vector $\boldsymbol{\theta}_{t,k}$ will have dimension $(3,1)$. 
\item For each pair $(i,t)$, the vector $\boldsymbol{\mu}_{i,t}$ will have dimension $(2,1)$. (Latent factors)
\item For each pair $(t,k)$, the vector $\boldsymbol{\lambda}_{t,k}$ will have dimension $(2,1)$
\end{itemize}



```{r, echo=FALSE}
# Set up chunk
library(MASS)
set.seed(215)

i = 3 # number of units
t = 3 # number of time points
k = 2 # number of outcomes
r = 2 # dimension of latent space.
```

\subsection{Generate $\delta_{t,k}$}

For $k\in 1,2$, want to generate
$$
\delta_{t,k} = \delta_{t-1,k} + \xi, \quad \xi\sim N(0,0.1), \delta_{1,k} \sim N(1,1).
$$
```{r, echo=FALSE}
dtk_mat <- matrix(NA, 3, 2)

# generate for k = 1, 2
initial_values <- rnorm(2, 1, 1)
# k = 1
d11 <- initial_values[1]
# k = 2
d12 <- initial_values[2] 

# now create the t = 2,3
for (i in 1:3) {
  prev <- if (i == 1) c(d11, d12) else dtk_mat[i-1, ]
  dtk_mat[i, ] <- prev + rnorm(2,0,0.1)
}

dtk_mat <- t(dtk_mat)
cat("Unit independent time trend  (for each outcome): \n")
dtk_mat
cat("\n")
```
First row is outcome $k=1$, second is $k=2$. Note: this adds the random noise that I left out of the DGP, due to $\xi$.

\subsection{Generating the $\mathbf{Z}_{i,t}$}

Let $\mathbf{Z}_{1,1} = (\varphi_1, \varphi_2,\varphi_3)$, and $\zeta_1 = (\zeta_{11}, \zeta_{12},\zeta_{13})$. We set:
$$
\mathbf{Z}_{1,2} = \mathbf{Z}_{1,1} + \zeta_1,
$$ and 
$$
\mathbf{Z}_{1,3} = \mathbf{Z}_{1,1} + 2 \zeta_1,
$$
to give time trends to the covariates.
```{r, echo=FALSE}
size <- 3
mean = c(0,0,0)
cov = matrix(c(1,0,0,0,1,0,0,0,1),ncol = 3) # identity matrix

# generating initial covariate measurements
Z <- mvrnorm(n = 3, mu = mean, Sigma = cov) 

# each row represents initial covariate measurements at t = 1, for each i in 1,2,3
z_1 <- Z[1,]
z_2 <- Z[2,]
z_3 <- Z[3,]

Zeta <- mvrnorm(n = 3, mu = mean, Sigma = cov) # use this to determine how covariates change over time.
Zeta_1 <- Zeta[1,]
Zeta_2 <- Zeta[2,]
Zeta_3 <- Zeta[3,]

make_Z_matrix <- function(z, zeta, T=3) {
  mat <- matrix(NA, nrow=T, ncol=length(z))
  for (i in 1:T) {
    mat[i,] <- z + ((i-1) * zeta)
  }
  mat
}

# columns of Z_1matrix, eg, represent a covariate changing over time.
# rows of Z_1matrix represent Z_{1,t} for t in 1,2,3
Z_1matrix <- make_Z_matrix(z_1, Zeta_1)
Z_2matrix <- make_Z_matrix(z_2, Zeta_2)
Z_3matrix <- make_Z_matrix(z_3, Zeta_3)

cat("Covariates for unit 1: \n")
Z_1matrix
cat("\n")

cat("Covariates for unit 2: \n")
Z_2matrix
cat("\n")

cat("Covariates for unit 2: \n")
Z_2matrix
cat("\n")
```
\subsection{Generate the $\boldsymbol{\mu}_{i,t}$}

We use a similar procedure to the generation of the observed covariates. For instance, the vector $\boldsymbol{\mu}_{1,1}$ has dimension 2, and we add a latent time trend in each latent factor through the repeated addition of $\zeta_{1}*$ (star to differentiate from $\zeta_1$). The three matrices we end up with correspond to the three units in the simulation.

```{r,echo=FALSE}
# Latent factors: dimension 2, independent of Z
mu_mean <- c(0, 0)
mu_cov  <- matrix(c(1,0,0,1), ncol = 2)

# Initial factor values for each unit
M <- mvrnorm(n = 3, mu = mu_mean, Sigma = mu_cov)
m_1 <- M[1,]
m_2 <- M[2,]
m_3 <- M[3,]

# Factor drift over time (analogous to Zeta)
Xi <- mvrnorm(n = 3, mu = mu_mean, Sigma = mu_cov)
Xi_1 <- Xi[1,]
Xi_2 <- Xi[2,]
Xi_3 <- Xi[3,]

# Build time-varying factor matrices (analogous to make_Z_matrix)
M_1matrix <- make_Z_matrix(m_1, Xi_1)
M_2matrix <- make_Z_matrix(m_2, Xi_2)
M_3matrix <- make_Z_matrix(m_3, Xi_3)
M_1matrix
```
\subsection{Generate $\boldsymbol{\theta}_{t,k}$ and $\boldsymbol{\lambda}_{t,k}$}
Create the 6 $(t\times k)$ loading vectors $\boldsymbol{\theta}_{t,k}$ each with dimension $(3,1)$.

Create the 6 $(t\times k)$ loading vectors for $\boldsymbol{\lambda}_{t,k}$ each with dimension $(2,1)$.
```{r, echo=FALSE}
theta <- matrix(NA, nrow = 6, ncol = 3)
for (i in 1:6) {
  # for binomial (0,1) loadings
  #theta[i,] <- rbinom(n=3, size = 1, p = 0.8)
  theta[i,] <- rnorm(n=3, mean=0, sd=1)
}

# rows of theta are the pairs of t,k.
cat("Loadings for observed covariates: \n")
theta
cat("\n")

lambda <- matrix(NA, nrow = 6, ncol = 2)
for (i in 1:6) {
  # for binomial (0,1) loadings
  # lambda[i,] <- rbinom(n=2, size = 1, p = 0.8)
  lambda[i,] <- rnorm(n=2, mean=0, sd=1)
}

# rows of lambda correspond to pairs of t,k
cat("Loadings for unobserved latent factors: \n")
lambda
cat("\n")
```

\subsection{Combining Factors and Loadings}
```{r, echo=FALSE}
rows = 6
cols = 3

# just a function to do special multiplication of factors times loadings.
load_factors <- function(fact_mat, load_mat) {
  output <- numeric(0)
  fact_mat <- rbind(fact_mat, fact_mat)
  for (i in 1:rows) {
    # k = 1, t = 1,2,3, then k = 2, t = 1,2,3
    output[i] <- sum(fact_mat[i,] * load_mat[i,])
  }
  # first row is k = 1
  # second row is k = 2
  output <- matrix(output, nrow = 2, ncol = cols)
  return (output)
}


Z_1loaded <- load_factors(Z_1matrix, theta)
Z_2loaded <- load_factors(Z_2matrix, theta)
Z_3loaded <- load_factors(Z_3matrix, theta)

cat("Unit 1 covariates times loadings: \n")
Z_1loaded
cat("\n")

cat("Unit 2 covariates times loadings: \n")
Z_2loaded
cat("\n")

cat("Unit 3 covariates times loadings: \n")
Z_3loaded
cat("\n")
```


```{r,echo=FALSE}


M_1loaded <- load_factors(M_1matrix, lambda)
M_2loaded <- load_factors(M_2matrix, lambda)
M_3loaded <- load_factors(M_3matrix, lambda)

cat("########################################")

cat("Unit 1 latent factors times loadings: \n")
M_1loaded
cat("\n")

cat("Unit 2 latent factors times loadings: \n")
M_2loaded
cat("\n")

cat("Unit 3 latent factors times loadings: \n")
M_3loaded
cat("\n")

```
\subsection{Adding it All Up}
```{r,echo=FALSE}

outcomes_unit1 <- dtk_mat + Z_1loaded + M_1loaded
outcomes_unit2 <- dtk_mat + Z_2loaded + M_2loaded
outcomes_unit3 <- dtk_mat + Z_3loaded + M_3loaded

cat("Outcomes for unit 1: \n")
outcomes_unit1
cat("\n")

cat("Outcomes for unit 2: \n")
outcomes_unit2
cat("\n")

cat("Outcomes for unit 3: \n")
outcomes_unit3
cat("\n")
```
```{r,echo=FALSE}
outcomes <- rbind(outcomes_unit1,outcomes_unit2, outcomes_unit3)

saveRDS(outcomes, "outcomes.rds")
```