---
title: "McSimDatGen"
output:
  pdf_document: default
  html_document: default
---

\textcolor{red}{Need to rename Z to X to be consistent across my different write ups.}
Goal is to generate outcomes using the model
$$
Y_{it,k}^0 = \delta_{t,k} + \mathbf{Z}_{i,t}' \boldsymbol{\theta}_{t,k} + \boldsymbol{\mu}_{i,t}\boldsymbol{\lambda}_{t,k} ,\quad\text{leaving out contribution from }\varepsilon_{itk}.
$$
In particular:
\begin{itemize}
\item For each pair $(t,k)$, $\delta_{t,k}$ is a scalar. 
\item For each pair $(i,t)$, the vector $\mathbf{Z}_{i,t}$ will have dimension $(3,1)$. (Observed covariate measurements)
\item For each pair $(t,k)$, the vector $\boldsymbol{\theta}_{t,k}$ will have dimension $(3,1)$. 
\item For each pair $(i,t)$, the vector $\boldsymbol{\mu}_{i,t}$ will have dimension $(2,1)$. (Latent factors)
\item For each pair $(t,k)$, the vector $\boldsymbol{\lambda}_{t,k}$ will have dimension $(2,1)$
\end{itemize}



```{r}
# Set up chunk
library(MASS)
set.seed(215)

i = 3 # number of units
t = 3 # number of time points
k = 2 # number of outcomes
r = 2 # dimension of latent space.
```

Generate delta t,k

For $k\in 1,2$, want to generate
$$
\delta_{t,k} = \delta_{t-1,k} + \xi, \quad \xi\sim N(0,0.1), \delta_{1,k} \sim N(1,1).
$$
```{r}
dtk_mat <- matrix(NA, 3, 2)

# generate for k = 1
initial_values <- rnorm(2, 1, 1)
d11 <- initial_values[1]
d12 <- initial_values[2] # good here

for (i in 1:3) {
  prev <- if (i == 1) c(d11, d12) else dtk_mat[i-1, ]
  dtk_mat[i, ] <- prev + rnorm(2,0,0.1)
}

# each of these represent the values of dtk for some k in 1,2.
dt1 <- dtk_mat[,1]
dt2 <- dtk_mat[,2]
```
Note: this adds the random noise that I left out of the DGP, due to $\xi$.

Generating the $\mathbf{Z}_{i,t}$.

Let $\mathbf{Z}_{1,1} = (\varphi_1, \varphi_2,\varphi_3)$, and $\zeta_1 = (\zeta_{11}, \zeta_{12},\zeta_{13})$. We set:
$$
Z_{1,2} = Z_{1,1} + \zeta_1,
$$ and 
$$
Z_{1,3} = Z_{1,1} + 2 \zeta_1,
$$
to give time trends to the covariates.
```{r}
# for each pair (i,t) there are two latent "observations"
# because we have chosen the dimension of the latent space to be 2
# this will require generating 9 vectors of dimension (2,1)

size <- 3
mean = c(0,0,0)
cov = matrix(c(1,0,0,0,1,0,0,0,1),ncol = 3) # identity matrix

# generating initial covariate measurements
Z <- mvrnorm(n = 3, mu = mean, Sigma = cov) 

# each row represents initial covariate measurements at t = 1, for each i in 1,2,3
z_1 <- Z[1,]
z_2 <- Z[2,]
z_3 <- Z[3,]

Zeta <- mvrnorm(n = 3, mu = mean, Sigma = cov) # use this to determine how covariates change over time.
Zeta_1 <- Zeta[1,]
Zeta_2 <- Zeta[2,]
Zeta_3 <- Zeta[3,]






# I'm creating the Z_it here.

```

Create the 6 $(t\times k)$ loading vectors $\boldsymbol{\theta}_{t,k}$ each with dimension $(3,1)$.
Create the 6 $(t\times k)$ loading vectors for $\boldsymbol{\lambda}_{t,k}$ each with dimension $(2,1)$.
```{r}
set.seed(215) 

theta <- matrix(NA, nrow = 6, ncol = 3)
for (i in 1:6) {
  theta[i,] <- rbinom(n=3, size = 1, p = 0.8)
}

# rows of theta are the pairs of t,k.
cat("Loadings for observed covariates: \n")
theta
cat("\n")

lambda <- matrix(NA, nrow = 6, ncol = 2)
for (i in 1:6) {
  lambda[i,] <- rbinom(n=2, size = 1, p = 0.8)
}

# rows of lambda correspond to pairs of t,k
cat("Loadings for unobserved latent factors: \n")
lambda
cat("\n")
```

Generate factor loadings
```{r}
size_phi <- 3 # number of samples we want
mean_vec_phi <- c(0,0)
cov_phi <- matrix(c(1,0,0,1),ncol=2) # identity matrix

phi_sample <- mvrnorm(n = size_phi,
                    mu = mean_vec_phi, 
                    Sigma = cov_phi) 

phi_sample

#phi_sample[2,] # first set of factor loadings
```
Goal now is to generate the $\boldsymbol{\mu}_{i,t}$ according to:
$$
\mu_{i,t} = t \mu_{i,1}\quad\text{with } \mu_{i,1}\sim N(0,1)
$$
for each $i\in 1,2,3$. \textcolor{red}{Is it a problem that there's no relationship between the latent factors and the observed covariates?}



Generate etas
```{r}
# generate six etas, since t x k = 6
size_eta <- 6 # number of observations of 
mean_vec_eta <- c(0,0)
cov_eta <- matrix(c(1,0,0,1),ncol=2) # identity matrix

eta_sample <- mvrnorm(n = size_eta,
                    mu = mean_vec_eta, 
                    Sigma = cov_eta) 

```
Generate lambdas from etas
```{r}
rho <- 1
lambda_1_mat <- matrix(NA, nrow = 3, ncol = k)
for (i in 1:3) {
  prev <- if (i == 1) rep(0, k) else lambda_1_mat[i-1, ]
  lambda_1_mat[i, ] <- rho * prev + eta_sample[i, ]
}

#mu_1_mat[3,] - mu_1_mat[2,] # check this worked. It did

lambda_2_mat <- matrix(NA, nrow = 3, ncol = k)
for (j in 1:3) {
  prev <- if (j == 1) rep(0, k) else lambda_2_mat[j-1, ]
  lambda_2_mat[j, ] <- rho * prev + eta_sample[j+3, ]
}

lambda_2_mat[3,] - lambda_2_mat[2,] # check this worked. It did



cat("lambda_1_mat: \n")
lambda_1_mat
cat("\n")

cat("lambda_2_mat: \n")
lambda_2_mat
cat("\n")
```

Generate alphas, betas
```{r}
# not using any errors.
sigma_alpha <- sqrt(0.5)
sigma_beta <- 1

ik <- 6
tk <- 6
itk <- 18

alpha_ik <- rnorm(ik, mean = 0, sd = sigma_alpha)
alpha_ik <- matrix(alpha_ik, nrow = 3, ncol = 2)

beta_tk <- rnorm(tk, mean = 0, sd = sigma_beta)
beta_tk <- matrix(beta_tk, nrow = 3, ncol = 2)

alpha_ik
beta_tk

```
Generate L matrix (product of factors and loadings)
```{r}
# want to do this in a vectorized way eventually
# k= 1
L_mat_1 <- matrix(NA, nrow = 3, ncol = 3)

for (i in 1:3) {
  for (j in 1:3) {
    L_mat_1[i,j] = sum(phi_sample[i,] * mu_1_mat[j,])
  }
}

# k = 2
L_mat_2 <- matrix(NA, nrow = 3, ncol = 3)

for (i in 1:3) {
  for (j in 1:3) {
    L_mat_2[i,j] = sum(phi_sample[i,] * mu_2_mat[j,])
  }
}


# at the end do a row bind step to produce final matrix
L_mat <- rbind(L_mat_1, L_mat_2)
L_mat # these are the 18 factor loadings.
```

Generate Ys
```{r}
# How are the covariates related to the outcomes? 

```


