---
title: "McSimDatGen"
output:
  pdf_document: default
  html_document: default
---
\section{Generating Data}

Goal is to generate outcomes using the model
$$
Y_{it,k}^0 = \delta_{t,k} + \mathbf{Z}_{i,t}' \boldsymbol{\theta}_{k} + \boldsymbol{\mu}_{i,t}\boldsymbol{\lambda}_{k} ,\quad\text{leaving out contribution from }\varepsilon_{itk}.
$$ which has time invariant factor loadings.
In particular:
\begin{itemize}
\item For each pair $(t,k)$, $\delta_{t,k}$ is a scalar. 
\item For each pair $(i,t)$, the vector $\mathbf{Z}_{i,t}$ will have dimension $(3,1)$. (Observed covariate measurements)
\item For each pair $(t,k)$, the vector $\boldsymbol{\theta}_{t,k}$ will have dimension $(3,1)$. 
\item For each pair $(i,t)$, the vector $\boldsymbol{\mu}_{i,t}$ will have dimension $(2,1)$. (Latent factors)
\item For each pair $(t,k)$, the vector $\boldsymbol{\lambda}_{t,k}$ will have dimension $(2,1)$
\end{itemize}



```{r, echo=FALSE}
# Set up chunk
library(MASS)
set.seed(215)


n_units = 4 # number of units (i)
n_timepoints = 10 # number of time points (t)
n_outcomes = 2 # number of outcomes (k)

t_0 = 5

n_predictors = 10 # number of predictors (length of each vector z_{i,t}) (p)
latent_dim = 4 # dimension of latent space. (l)

verbose = FALSE
```

\subsection{Generate $\delta_{t,k}$}

For $k\in 1,2$, want to generate
$$
\delta_{t,k} = \delta_{t-1,k} + \xi, \quad \xi\sim N(0,0.1), \delta_{1,k} \sim N(1,1).
$$
```{r, echo=FALSE}
# t = n_timepoints
# k = n_outcomes


# a function that creates the delta tk entries.
make_dtk <- function(n_timepoints,n_outcomes) {
  # initialize an empty matrix of dim t,k
  dtk_mat <- matrix(NA, n_timepoints, n_outcomes)
  # first row is random noise of length n_outcomes
  dtk_mat[1,] <- rnorm(n_outcomes, 1, 1)
  # fill in the remaining rows:
  for (i in 2:n_timepoints) {
    prev <- if (i==1) dtk_mat[1,] else dtk_mat[i-1,]
    dtk_mat[i,] <- prev + rnorm(n_outcomes,0,0.1)
  }
  return(dtk_mat)
}

dtk_matrix <- make_dtk(n_timepoints,n_outcomes) * 0.1

if (verbose == TRUE) {
  cat("dtk_matrix: \n")
  print(dtk_matrix)
  cat("\n")
}

```

First row is outcome $k=1$, second is $k=2$. Note: this adds the random noise that I left out of the DGP, due to $\xi$.

\subsection{Generating the $\mathbf{Z}_{i,t}$}

Let $\mathbf{Z}_{1,1} = (\varphi_1, \varphi_2,\varphi_3)$, and $\zeta_1 = (\zeta_{11}, \zeta_{12},\zeta_{13})$. We set:
$$
\mathbf{Z}_{1,2} = \mathbf{Z}_{1,1} + \zeta_1,
$$ and 
$$
\mathbf{Z}_{1,3} = \mathbf{Z}_{1,1} + 2 \zeta_1,
$$
to give time trends to the covariates.


```{r,echo=FALSE}
# number of predictors is dimension of covariate vectors. 
# Not necessarily the same as the length of the outcome vectors
mean = rep(0, n_predictors)
cov = diag(n_predictors) # identity matrix

# generating initial covariate measurements
Z <- mvrnorm(n = n_units, mu = mean, Sigma = cov) 
# each row represents initial covariate measurements at t = 1, for each i in 1,...,n_units
if (verbose == TRUE) {
  cat("Initial Covariate Measurements: \n")
  print(Z)
  cat("\n")
}


# Zeta determines how covariates change over time.
Zeta <- mvrnorm(n = n_units, mu = mean, Sigma = cov) #* 0.1 # to smooth time trends

# a function to apply Zeta to each row of the initial Z matrix
make_Z_matrix <- function(z, zeta, n_timepoints) {
  mat <- matrix(NA, nrow=n_timepoints, ncol=length(z))
  for (i in 1:n_timepoints) {
    mat[i,] <- z + ((i-1) * zeta)
  }
  return(mat)
}

# columns of Z_matrices[[1]], eg, represent a covariate changing over time.
Z_matrices <- lapply(1:n_units, function(i) {
  make_Z_matrix(Z[i,], Zeta[i,], n_timepoints = n_timepoints)
})

# print at most 5
if (verbose == TRUE) {
  for (i in 1:min(n_units,5)){
    cat("Covariates for unit ",i,"\n")
    print(Z_matrices[[i]])
    cat("\n")
  }
}

```

\subsection{Generate the $\boldsymbol{\mu}_{i,t}$}

We use a similar procedure to the generation of the observed covariates. For instance, the vector $\boldsymbol{\mu}_{1,1}$ has dimension 2, and we add a latent time trend in each latent factor through the repeated addition of $\zeta_{1}*$ (star to differentiate from $\zeta_1$). The three matrices we end up with correspond to the three units in the simulation.

```{r,echo=FALSE}
mu_mean <- rep(0, latent_dim)
mu_cov  <- diag(latent_dim)

# Initial factor values for each unit
M <- mvrnorm(n = n_units, mu = mu_mean, Sigma = mu_cov)

# Factor drift over time (analogous to Zeta)
Xi <- mvrnorm(n = n_units, mu = mu_mean, Sigma = mu_cov) #*0.1 # to smooth time trends

if (verbose == TRUE) {
  cat("Initial Factor Values: \n")
  print(M)
  cat("\n")
  
  cat("Factor Drift: \n")
  print(Xi)
  cat("\n")
}

M_matrices <- lapply(1:n_units, function(i) {
  make_Z_matrix(M[i,], Xi[i,], n_timepoints = n_timepoints)
})

if (verbose == TRUE) {
  cat("M_matrices: \n")
  print(M_matrices)
  cat("\n")
}


```


\subsection{Generate $\boldsymbol{\theta}_{k}$ and $\boldsymbol{\lambda}_{k}$}
Create the $k$ loading vectors $\boldsymbol{\theta}_{k}$ each with dimension $(p,1)$.

Create the $k$ loading vectors for $\boldsymbol{\lambda}_{k}$ each with dimension $(\ell,1)$.
```{r, echo=FALSE}

# generate time invariant observed covariate loadings
theta <- matrix(NA, nrow = n_outcomes, ncol = n_predictors)
for (i in 1:n_outcomes) {
  theta[i,] <- rnorm(n=n_predictors, mean=0, sd=0.1/sqrt(n_predictors))
}


if (verbose==TRUE) {
  cat("Loadings for observed covariates: \n")
  cat("(One row is one theta_k vector) \n")
  print(theta)
  cat("\n")
}

# generate time invariant latent factor loadings
lambda <- matrix(NA, nrow = n_outcomes, ncol = latent_dim)
for (i in 1:n_outcomes) {
  lambda[i,] <- rnorm(n=latent_dim, mean=0, sd=0.1/sqrt(latent_dim))
}

if (verbose==TRUE) {
  cat("Loadings for unobserved latent factors: \n")
  cat("(One row is one lambda_k vector) \n")
  print(lambda)
  cat("\n")
}

```

\subsection{Adding it all up}
```{r}
Y <- array(NA, dim = c(n_units, n_timepoints, n_outcomes))

for (i in 1:n_units) {
  for (k in 1:n_outcomes) {
    for (t in 1:n_timepoints) {
      delta_tk <- dtk_matrix[t,k]
      # take the product of n(0,1) * n(0,1) has variance 1. We add together n_predictors. 
      # so the variance of the product is n_predictors. We standardize when we are generating
      # the loadings /sqrt(n_predictors)
      z_theta_itk <- (Z_matrices[[i]][t,] %*% theta[k,]) 
      # also standardized the same way here
      mu_lambda_itk <- (M_matrices[[i]][t,] %*% lambda[k,]) 
      
      Y[i,t,k] <- delta_tk + z_theta_itk + mu_lambda_itk
    }
  }
}
verbose = TRUE
if (verbose==TRUE) {
  cat("The final data Y, by unit: \n")
  for (i in 1:n_units) {
    cat("Unit ",i," timepoints x outcomes\n")
    print(Y[i,,])
    cat("\n")
  }
}


# stack the matrices
Y_stacked <- do.call(rbind, lapply(1:n_timepoints, function(i) t(Y[,i,])))
cat("All unit outcomes stacked: \n")
print(Y_stacked)
cat("\n")

```

```{r}
write.csv(Y_stacked, "outcomes.csv", row.names = FALSE)

metadata <- c(n_units, n_timepoints, n_outcomes)
write.csv(metadata, "metadata.csv", row.names = c("n_units","n_timepoints","n_outcomes"))
```