---
title: "McSimDatGen"
output:
  pdf_document: default
  html_document: default
---
\section{Generating Data}

Goal is to generate outcomes using the model
$$
Y_{it,k}^0 = \delta_{t,k} + \mathbf{Z}_{i,t}' \boldsymbol{\theta}_{k} + \boldsymbol{\mu}_{i,t}\boldsymbol{\lambda}_{k} ,\quad\text{leaving out contribution from }\varepsilon_{itk}.
$$ which has time invariant factor loadings.
In particular:
\begin{itemize}
\item For each pair $(t,k)$, $\delta_{t,k}$ is a scalar. 
\item For each pair $(i,t)$, the vector $\mathbf{Z}_{i,t}$ will have dimension $(3,1)$. (Observed covariate measurements)
\item For each pair $(t,k)$, the vector $\boldsymbol{\theta}_{t,k}$ will have dimension $(3,1)$. 
\item For each pair $(i,t)$, the vector $\boldsymbol{\mu}_{i,t}$ will have dimension $(2,1)$. (Latent factors)
\item For each pair $(t,k)$, the vector $\boldsymbol{\lambda}_{t,k}$ will have dimension $(2,1)$
\end{itemize}


\subsection{Generate $\delta_{t,k}$}

For $k \in \{1, \ldots, \texttt{n\_outcomes}\}$, want to generate
$$
\delta_{t,k} = \delta_{t-1,k} + \xi, \quad \xi\sim N(0,0.1), \delta_{1,k} \sim N(1,1).
$$
First row is outcome $k=1$, second is $k=2$, etc. Note: I believe this adds the random noise that I left out of the DGP, due to $\xi$.

\subsection{Generating the $\mathbf{Z}_{i,t}$}

Let $\mathbf{Z}_{1,1} = (\varphi_1, \varphi_2,\varphi_3)$, and $\zeta_1 = (\zeta_{11}, \zeta_{12},\zeta_{13})$. We set:
$$
\mathbf{Z}_{1,2} = \mathbf{Z}_{1,1} + \zeta_1,
$$ and 
$$
\mathbf{Z}_{1,3} = \mathbf{Z}_{1,1} + 2 \zeta_1,
$$
to give time trends to the covariates.

\subsection{Generate the $\boldsymbol{\mu}_{i,t}$}

We use a similar procedure to the generation of the observed covariates. For instance, the vector $\boldsymbol{\mu}_{1,1}$ has dimension 2, and we add a latent time trend in each latent factor through the repeated addition of $\zeta_{1}*$ (star to differentiate from $\zeta_1$). The three matrices we end up with correspond to the three units in the simulation.

\subsection{Generate $\boldsymbol{\theta}_{k}$ and $\boldsymbol{\lambda}_{k}$}
Create the $k$ loading vectors $\boldsymbol{\theta}_{k}$ each with dimension $(p,1)$.

Create the $k$ loading vectors for $\boldsymbol{\lambda}_{k}$ each with dimension $(\ell,1)$.

```{r, echo=FALSE}
# Set up
library(MASS)
set.seed(215)

### ------------------------------------------------------------ ###

# t = n_timepoints
# k = n_outcomes

# a function that creates the delta tk entries.
make_dtk <- function(n_timepoints, n_outcomes, treatment_effect, treatment_time) {
  # initialize an empty matrix of dim t,k
  dtk_mat <- matrix(NA, n_timepoints, n_outcomes)
  # first row is random noise of length n_outcomes
  dtk_mat[1,] <- rnorm(n_outcomes, 1, 1)
  # fill in the remaining rows:
  for (i in 2:n_timepoints) {
    if (treatment_effect==TRUE & i > treatment_time) {
      effect <- rnorm(n_outcomes,5,0.1)
    } else {
      effect <- rnorm(n_outcomes,0,0.1)
    }
    prev <- if (i==1) dtk_mat[1,] else dtk_mat[i-1,]
    dtk_mat[i,] <- prev + effect
  }
  return(dtk_mat)
}

### ------------------------------------------------------------ ###

make_Z_matrix <- function(z, zeta, n_timepoints) {
  z = z
  zeta = zeta
  n_timepoints = n_timepoints
  
  mat <- matrix(NA, nrow=n_timepoints, ncol=length(z))
  for (i in 1:n_timepoints) {
    mat[i,] <- z + ((i-1) * zeta)
  }
  return(mat)
}

### ------------------------------------------------------------ ###

# a function to make the time varying covariates
# NB: number of predictors is dimension of covariate vectors. 
# Not necessarily the same as the length of the outcome vectors
make_Z_matrices <- function(n_predictors, n_timepoints) {
  n_predictors = n_predictors
  n_timepoints = n_timepoints
  
  mean = rep(0, n_predictors)
  cov = diag(n_predictors) # identity matrix
  
  # generating initial covariate measurements
  # each row represents initial covariate measurements at t = 1, for each i in 1,...,n_units
  Z <- mvrnorm(n = n_units, mu = mean, Sigma = cov) 
  # Zeta determines how covariates change over time.
  Zeta <- mvrnorm(n = n_units, mu = mean, Sigma = cov) #* 0.1 # to smooth time trends
  
  # a function to apply Zeta to each row of the initial Z matrix
  
  # columns of Z_matrices[[1]], eg, represent a covariate changing over time.
  Z_matrices <- lapply(1:n_units, function(i) {
    make_Z_matrix(Z[i,], Zeta[i,], n_timepoints = n_timepoints)
  })
  return(Z_matrices)
}

### ------------------------------------------------------------ ###

# makes the time varying latent factors.
make_M_matrices <- function(latent_dim, n_timepoints) {
  latent_dim = latent_dim
  n_timepoints = n_timepoints
  
  mu_mean <- rep(0, latent_dim)
  mu_cov  <- diag(latent_dim)
  
  # Initial factor values for each unit
  M <- mvrnorm(n = n_units, mu = mu_mean, Sigma = mu_cov)
  
  # Factor drift over time (analogous to Zeta)
  Xi <- mvrnorm(n = n_units, mu = mu_mean, Sigma = mu_cov) #*0.1 # to smooth time trends
  
  M_matrices <- lapply(1:n_units, function(i) {
    make_Z_matrix(M[i,], Xi[i,], n_timepoints = n_timepoints)
  })
  return(M_matrices)
}

### ------------------------------------------------------------ ###

# make the time invariant covariate loadings
make_theta <- function(n_outcomes, n_predictors) {
  n_outcomes = n_outcomes
  n_predictors = n_predictors
  
  theta <- matrix(NA, nrow = n_outcomes, ncol = n_predictors)
  for (i in 1:n_outcomes) {
    theta[i,] <- rnorm(n=n_predictors, mean=0, sd=0.1/sqrt(n_predictors))
  }
  if (verbose==TRUE) {
    cat("Loadings for observed covariates: \n")
    cat("(One row is one theta_k vector) \n")
    print(theta)
    cat("\n")
  }
  return (theta)
}

### ------------------------------------------------------------ ###

# generate time invariant latent factor loadings
make_lambda <- function(n_outcomes, latent_dim) {
  n_outcomes = n_outcomes
  latent_dim = latent_dim
  
  lambda <- matrix(NA, nrow = n_outcomes, ncol = latent_dim)
  for (i in 1:n_outcomes) {
    lambda[i,] <- rnorm(n=latent_dim, mean=0, sd=0.1/sqrt(latent_dim))
  }
  
  if (verbose==TRUE) {
    cat("Loadings for unobserved latent factors: \n")
    cat("(One row is one lambda_k vector) \n")
    print(lambda)
    cat("\n")
  }
  return(lambda)
}

### ------------------------------------------------------------ ###

# standardize per outcome using only the pretreatment info.
scale_function <- function(Y,n_units, n_timepoints,n_outcomes,treatment_time) {
  for (i in 1:n_outcomes) {
    mat <- matrix(Y[,,i], nrow=n_units, ncol = n_timepoints)
    cat("pre-treatment sub-matrix: \n")
    observable_mat <- mat[,1:treatment_time]
    print(observable_mat)
  
    # calculate the mean and sd of the observable data matrix
    tot_mean <- mean(observable_mat)
    vect <- as.vector(observable_mat)
    tot_sd <- sd(vect)
    
    mat <- (mat - tot_mean) / tot_sd
    Y[,,i] <- mat
  }
  Y_scaled <- Y
  return (Y_scaled)
}

### ------------------------------------------------------------ ###

# notes for (n_units, n_timepoints)
# (4, 10) gives good similarity (4,20) is even better
# (10, 10) gives great similarity (10, 20) is also extremely good
# turning up n_outcomes to 3 it still works great with (10,20) and (10,10)
# (8,8) embedding into 9 is a good example to show jing

# inputs to the generate data function
n_units = 8 # number of units (i)
n_timepoints = 8 # number of time points (t)
n_outcomes = 3 # number of outcomes (k)

n_predictors = 10 # number of predictors (length of each vector z_{i,t}) (p)
latent_dim = 4 # dimension of latent space. (l)

treatment_effect = TRUE # should we generate the data with a treatment effect
treatment_time = 5 # must be less than n_timepoints

verbose = FALSE 

### ------------------------------------------------------------ ###

# generates the synthetic data. Calls several other functions, above
generate_data <- function(n_units, n_timepoints, n_outcomes, n_predictors, latent_dim, treatment_effect, treatment_time, verbose) {
  dtk_matrix <- make_dtk(n_timepoints,n_outcomes, treatment_effect, treatment_time) * 0.1
  if (verbose == TRUE) {
    cat("dtk_matrix: \n")
    print(dtk_matrix)
    cat("\n")
  }
  
  Z_matrices <- make_Z_matrices(n_predictors, n_timepoints)
  M_matrices <- make_M_matrices(latent_dim, n_timepoints)
  theta <- make_theta(n_outcomes, n_predictors)
  lambda <- make_lambda(n_outcomes, latent_dim)
  
  # an array to store the synthetic outcomes
  Y <- array(NA, dim = c(n_units, n_timepoints, n_outcomes))
  
  # add everything together
  for (i in 1:n_units) {
    for (k in 1:n_outcomes) {
      for (t in 1:n_timepoints) {
        delta_tk <- dtk_matrix[t,k]
        # take the product of n(0,1) * n(0,1) has variance 1. We add together n_predictors. 
        # so the variance of the product is n_predictors. We standardize when we are generating
        # the loadings /sqrt(n_predictors)
        z_theta_itk <- (Z_matrices[[i]][t,] %*% theta[k,]) 
        # also standardized the same way here
        mu_lambda_itk <- (M_matrices[[i]][t,] %*% lambda[k,]) 
        
        Y[i,t,k] <- delta_tk + z_theta_itk + mu_lambda_itk
      }
    }
  }
  verbose = FALSE
  if (verbose==TRUE) {
    cat("The final data Y, by unit: \n")
    for (i in 1:n_units) {
      cat("Unit ",i," timepoints x outcomes\n")
      print(Y[i,,])
      cat("\n")
    }
  }
  
  
  
  Y <- scale_function(Y,n_units, n_timepoints,n_outcomes,treatment_time)
  
  # stack the matrices                                    
  Y_final <- do.call(rbind, lapply(1:n_timepoints, function(i) t(Y[,i,])))
  cat("All unit outcomes stacked: \n")
  print(Y_final)
  cat("\n")
  
  # output information
  write.csv(Y_final, "outcomes.csv", row.names = FALSE)
  metadata <- c(n_units, n_timepoints, n_outcomes)
  write.csv(metadata, "metadata.csv", row.names = c("n_units","n_timepoints","n_outcomes"))
}

### ------------------------------------------------------------ ###

# call the function
generate_data(n_units, n_timepoints, n_outcomes, n_predictors, latent_dim, treatment_effect, treatment_time, verbose)


```




```{r}

# Staging ground for standardizing.

# need to normalize using only pre-treatment - first t_0 rows
set.seed(215)

n_units = 3
n_timepoints = 3
n_outcomes = 2
treatment_time = 2

# Creates a 3x4x2 array of random integers between 1 and 100
X <- array(sample(1:100, 24, replace = TRUE), dim = c(n_units, n_timepoints, n_outcomes))
print(X)

for (i in 1:n_outcomes) {
  mat <- matrix(X[,,i], nrow=n_units, ncol = n_timepoints)
  cat("pre-treatment sub-matrix: \n")
  observable_mat <- mat[,1:treatment_time]
  print(observable_mat)

  # calculate the mean and sd of the observable data matrix
  tot_mean <- mean(observable_mat)
  vect <- as.vector(observable_mat)
  tot_sd <- sd(vect)
  
  mat <- (mat - tot_mean) / tot_sd
  X[,,i] <- mat
}
```






