---
title: "replicateEBMcode"
output:
  pdf_document: default
  html_document: default
---

Loads packages and removes loaded data
```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(MASS)
rm(list = ls())
```

Color palate, figure names, rho, and create a directory
```{r}
# The palette with black:
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# Change the weights in model to reflect importance of common factors
rho <- 1
bias_fig_name <- 'figure/all_bias_sims_n50_rho1.pdf'
imbalance_fig_name <- 'figure/all_imbalance_sims_n50_rho1.pdf'
dir.create("figureJS", showWarnings = FALSE)
```

In R, "osqp" solves quadratic programs of the form
$$
\min \frac{1}{2} x^TPx + q^Tx \quad\text{subject to}\quad l\leq Ax\leq u
$$
Let $\mathbf{X}_0$ be the matrix of control outcomes, and $\vec{X}_1$ be the target vector. The synth_qp function solves the problem:
$$
\min_{x} x^T \mathbf{X}_0 ^T \mathbf{X}_0 x - 2 (\vec{X}_1^T \mathbf{X}_0) x \quad\text{subject to}\quad \sum_{i} x_i = 1\quad\text{and}\quad 1 \geq x_i \geq 0,
$$ which is precisely the QP:
$$
\min_{x}|| \vec{X}_1 - \mathbf{X}_0 x|| \quad\text{subject to}\quad \sum_i x_i = 1, \quad\text{and}\quad 1 \geq x_i \geq 0.
$$

In the EBM code, the QP being described is
```{r}
#' Solve the synth QP directly
#' @param X1 Target vector
#' @param X0 Matrix of control outcomes
#' @noRd
synth_qp <- function(X1, X0) {
  
  Pmat <- X0 %*% t(X0) # X^T X
  qvec <- - t(X1) %*% t(X0)
  
  n0 <- nrow(X0)
  A <- rbind(rep(1, n0), diag(n0))
  l <- c(1, numeric(n0))
  u <- c(1, rep(1, n0))
  
  settings = osqp::osqpSettings(verbose = FALSE,
                                eps_rel = 1e-8,
                                eps_abs = 1e-8)
  sol <- osqp::solve_osqp(P = Pmat, q = qvec,
                          A = A, l = l, u = u, 
                          pars = settings)
  
  return(sol$x)
}
```
And the matrix $A$, along with the vectors $l$ and $u$ are encoding the constraints described above.

\textcolor{red}{Come back to this calculation. What is it doing?}
The first entry is covariates, second is target, third is weights. This returns the bias of the SCM weights.
```{r}
##### Bias 
SCMbias <- function(Y0,Y1,weights) {
  synthY0 <- (Y0%*%weights)
  gap <- Y1-synthY0 
  return(gap)
}
```

## Generating the factor model data:

($\mu$ are factors, $\phi$ are factor loadings)

EBM generates data according to two factor models in the simulation study
\begin{itemize}
\item[(1)] A simpler model:
$$
Y_{itk}(0) = \phi_i \mu_{t} + \varepsilon_{itk},
$$ where multiple outcomes can be thought of as repeated (noisy) measurements of the same underlying quantity.
\item[(2)] A more complex model:
$$
Y_{itk}(0) = \rho\phi_i \mu_{t} + (1-\rho)\phi_{ik}\mu_{tk} + \varepsilon_{itk}
$$ to investigate how the "presence of idiosyncratic factors influence performance."
\end{itemize}

Create a vector $\mu$ of length $N$, which contains decreasing sequence.
This is for the simpler model.
```{r}
# factor loadings
n <- 3 # number of units, the first unit is treated

mu <- list(
  c = seq(5, 1, length.out = n) # N x 1 vector
)
mu

```
Switch the first two elements in $\mu$ and create a vector of length $n$ called trt which stores 1 in the first place, and all zeros otherwise. Represents that the treated unit is the first unit in $\mu$? 
```{r}
mu$c[c(1, 2)] <- mu$c[c(2, 1)];
trt <- numeric(n); trt[1] <- 1
trt
```

With a specific setting selected (choice of number of time points and outcomes), a vector factors is created which has length $T+1$. 
```{r}
t_total <- 4
k_total <- 3

factors <- list(
  c = seq(0.5, 1, length.out = t_total+1) # hand-code factor (common part)
)

factors
```

factors\$i holds the idiosyncratic part of the factors in a matrix of dimension $(T+1, K)$.
```{r}
factors$i <- matrix(0,nrow=t_total+1, ncol = k_total)
factors
```

Now we create independent AR series across the factors, ie across $\mu_{tk}$, so that as $t$ increases, $\mu_{tk}$ staggers upward. These are generated independently for each outcome $k$.
```{r}
set.seed(1)
for ( k in 1:k_total ) {
  series <- arima.sim(model = list(ar = 0.5), n = t_total+1) # independent AR series across outcomes
  factors$i[,k] <- 0.5 + (series - min(series)) * (1 - 0.5) / (max(series) - min(series))  # rescale so factors are in same range
  factors$i[,k] <- sort(factors$i[,k])
}

factors
```
Next, EBM creates the factor loadings $\phi_{ik}$ for the idiosyncratic component. These are randomly generated for each unit and each outcome.

For the more complex model. 

Note that mu\$i contains $ik$ - these are the factor loadings for the idiosyncratic component.

```{r}
set.seed(1)
mu$i <- matrix(0,nrow=n, ncol = k_total) # outcome idiosyncratic part of factor
for (k in (1:k_total)) {
  mu$i[,k] <- rnorm(n)
  mu$i[,k] <- 1 + mu$i[,k] * (5 - 1) / (max(mu$i[,k] ) - min(mu$i[,k] )) # rescale so bias is in same range
}

mu
```

Using the oracle weights, examining the bias that results
EBM considers two possibilities here. Denote the oracle weights by $\gamma^*$. 
The vector of outcomes for the treated unit is:
$$
Y_{1,T+1,k} = \phi_1\mu_{T+1}
$$
For the control units, the outcomes are calculated as:
$$
Y_{i,T+1,k} = \phi_i \mu_{T+1}\quad\text{for}\quad i \in 2,3.
$$
Then the bias is calculated as:
$$
Y_{1,T+1,k} - \sum_{i=1}^2 Y_{i,T+1,k} \gamma_i^*
$$ which is close to zero, since the weights are oracle weights.

A similar calculation is performed to ensure that the oracle weights are unbiased under the model with idiosyncratic factors.
```{r}
# just the entry containing the 
mu$c[trt]
mu$c[-trt]

theo_w <- synth_qp(mu$c[trt], as.matrix(mu$c[-trt]));
theo_w

# first entry is covariates, second is target, third is weights
SCMbias(mu$c[-trt]*factors$c[t_total+1], mu$c[trt]*factors$c[t_total+1],theo_w)

mu$i[trt,] <- t(theo_w)%*%mu$i[-trt,] # overwrite the idiosyncratic loadings to ensure oracle weights exist
  
theo_w <- synth_qp(c(mu$c[trt],mu$i[trt,]), as.matrix(cbind(mu$c[-trt],mu$i[-trt,]))); # confirm we can find the oracle weights
SCMbias(mu$c[-trt]*factors$c[t_total+1] + mu$i[-trt,1]*factors$i[t_total+1,1],
        mu$c[trt]*factors$c[t_total+1] + mu$i[trt,1]*factors$i[t_total+1,1], 
        theo_w)
```

```{r}
# create the model components
# first column of factors gets the non-AR process time series 
factors$i[,1] <- factors$c
factors
```
If $N=3$, create a matrix called "factors\_expanded" which contains:
$$
\begin{pmatrix}
\vec{\mu}_{1k}\\
\vec{\mu}_{1k}\\
\vec{\mu}_{1k} \\
\vdots \\
\vec{\mu}_{T+1k}\\
\vec{\mu}_{T+1k}\\
\vec{\mu}_{T+1k}
\end{pmatrix}
$$ 

```{r}
# Repeat each row of factors N times (expand to NT x K)
factors_expanded <- factors$i[rep(1:(t_total+1), each = n), ]
factors_expanded

```

```{r}
# put the simpler model loadings into the first column of loadings.
mu$i[,1] <- mu$c
mu
```

Create a copy of $\phi$ for each time point 1 through $T+1$.
```{r}
mu_expanded <- mu$i[rep(1:n, (t_total+1)), ]
mu_expanded

```

Perform the multiplication of $\phi_{ik}\mu_{tk}$.
```{r}
factors_expanded * mu_expanded
```

Final "model" encodes both the simpler model and the more complex model depending on the choice of $rho$. If we use the simpler model
```{r}
rho = 1
model <- rho * as.vector(factors$c %x% mu$c) + (1-rho) * factors_expanded * mu_expanded
cat("simpler model:\n")
model

rho = 0.9
model <- rho * as.vector(factors$c %x% mu$c) + (1-rho) * factors_expanded * mu_expanded
cat("complicated model (with AR process in factors):\n")
model
```
Once the data is generated, it is passed to a function called factor\_sim.

EBM begins by creating a random noise term, and adding it to the synthetic data
$$
\varepsilon_{itk} \sim N(0,1)
$$ (where I've coded the variance to be 1).

```{r}
n = 3
t_total <- 5
k_total <- 3
variance <- 1

epsilon <- matrix(rnorm((n * t_total) * k_total, sd = sqrt(variance)), ncol = k_total)
epsilon
```
Without the epsilon term, out is simply the data generated from the more complicated model:
```{r}
out <- model[1:(n * t_total),]
out
```

Adding the noise term, we have:
```{r}
out <- model[1:(n * t_total),] + epsilon
out
```
which represent noisy measurements of the outcomes.

EBM creates a binary mask to extract the time periods (rows of out) corresponding to the treated unit:
```{r}
t_trt <- rep(trt,t_total)
t_trt
```

```{r}
out_trt <- matrix(out[which(t_trt==1),],nrow = t_total,ncol =k_total)  # T x K matrix of outcomes for the treated unit

out_trt
```
Then EBM uses the same strategy (binary mask) to extract the time periods for the control units:
```{r}
out_control <- out[which(t_trt==0),]
out_control
```


The SCM weights are then calculated, according to each of the three methods:

First, a separate SC is fitted for each outcome
```{r}
## calculate SCM weights
out_trt_sep <- out_trt[,1]
out_control_sep <- matrix(out_control[,1], nrow = n-1,ncol=t_total)
w_sep <- synth_qp(out_trt_sep, out_control_sep);
r.svd <- svd(rbind(out_control_sep,out_trt_sep))
largest_svd <- r.svd$d[1]^2/sum(r.svd$d^2)
cond <-  r.svd$d[1]/r.svd$d[t_total]

w_sep
out_control_sep
``` 
which are the first outcome, for each control unit, and columns are time periods. The weights are calculated for 